Stack Safety as a Security Property

### 2020-07-06 main.tex
Some notes on notation in lazy properties

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machOT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
\end{minipage}

@@

@@

      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
      agrees with \(\mach_e\) on every component the callee should not change,
      but with \(\mach_r\) on those components that it is permitted to change.
      So the observable behavior that follows is the idealized behavior to
      which we compare the trace from \(\mach_r\), which must prefix it (to
      account for premature halts due to policy violations.)
      %
      \[\begin{split}
        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machOT' \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machOT')
      \end{split}\]

@@

    For out example, we return to the violation of trace integrity in
    \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
    Whether observable integrity holds depends on the continuing trace, shown
    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
    \(\mach_r\), with high integrity components in red and low integrity in
    green. Next is the rolled-back state, \(\mach_r' =
    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
    components are restored to their value as of \(\mach_e\). Specifically,
    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
    it prints 0. So observable trace integrity does not hold.

@@

      A system enjoys {\em observable integrity} with regard to \(\callmap\)
      and  \(\codemap\) if for any initial state \(\mach_0\) and each
      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
      contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

@@

      To return to the example in \cref{fig:confex}, the varied final state
      \(\mach_r'\) might do anything depending on the values of the variation,
      so we cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\). We must first restore the varied values to their original
      state, while preserving differences that have occurred during
      execution.

@@

      A pair of traces \(\MPT\) and \(\machOT\) starting from \(\mach_e\) and
      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\),
      enjoy observable trace confidentiality with respect to \(\contour\) if
      two conditions hold:

@@

      If \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machOT\) is finite. Let
      \(\mach_r'\) be the final state of \(\machOT\). Then the induced traces
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
      \hookrightarrow \machOT'\) are observationally prefixed,
      \(\pi_o(\MPT') \lesssim \pi_o(\machOT')\).

      We write this \(\text{\sc ObsConf}_\contour ~ \MPT\).

@@

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
      \mach_e,\mach_e',\mach_r,\mach_r'\). Addresses 108 and 109, in pink, are
      not restored; every address that was varied and not changed in one of the
      executions is. Since 108 contains a value that originates in the
      variation, values that depend on it might be anything; after returning
      from {\tt g}, {\tt main} prints 95 in the trace from \(\mach_r\) and
      some undetermined value in the trace from \(\mach_r''\). So observable
      trace confidentiality does not hold.

@@

    \definition{Observable Stack Safety}
      A system enjoys {\em observable stack safety} with regard to a call map
      and a code map if it enjoys both observable integrity and observable
      confidentiality with regard to them.

>>>>

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machOT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
\end{minipage}
\rb{Shouldn't the conclusion of the first judgment use $\hookrightarrow_P$?
  And in the second judgment, wouldn't $MO$ be $MPO$, and the third premise $\hookrightarrow_P$?}

@@

    \rb{Suggestion: name this ``observable \emph{trace? something else?}
      integrity'' to avoid confusion with observable stack integrity (in fact,
      it is called exactly this when discussing the example below). The rollback
      may also be a definition of its own to match the granularity of
      definitions in previous sections.}

@@

      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
      agrees with \(\mach_e\) on every component the callee should not change,
      but with \(\mach_r\) on those components that it is permitted to change.
      So the observable behavior that follows is the idealized behavior to
      which we compare the trace from \(\mach_r\), which must prefix it (to
      account for premature halts due to policy violations.)
      \rb{Here, should the second premise use $\hookrightarrow_P$?}
      %
      \[\begin{split}
        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machOT' \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machOT')
      \end{split}\]

@@

    For out example, we return to the violation of trace integrity in
    \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
    Whether observable integrity holds depends on the continuing trace, shown
    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
    \(\mach_r\), with high integrity components in red and low integrity in
    green. Next is the rolled-back state, \(\mach_r' =
    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
    components are restored to their values as of \(\mach_e\). Specifically,
    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
    it prints 0. So observable trace integrity does not hold.

@@

      A system enjoys {\em observable \rb{stack?} integrity} with regard to \(\callmap\)
      and  \(\codemap\) if for any initial state \(\mach_0\) and each
      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
      contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\) \rb{$\hookrightarrow_P$?},
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

@@

      To return to the example in \cref{fig:confex}, the varied final state
      \(\mach_r'\) might do anything depending on the values of the variation,
      so we cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\). We must first restore the varied values to their original
      state, while preserving differences that have occurred during
      execution.
      \rb{To me, this reads like a prologue to the definitions properly
        speaking---maybe the definition header above should be a paragraph and
        a definition header appear next? I would also split it into two
        definitions.}

@@

      A pair of traces \(\MPT\) and \(\machOT\) starting from \(\mach_e\) \rb{with what $\pol_e$?} and
      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\) \rb{$\mach_e'$?},
      enjoy observable trace confidentiality with respect to \(\contour\) if
      two conditions hold:

@@

      If \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machOT\) is finite. Let
      \(\mach_r'\) be the final state of \(\machOT\). Then the induced traces
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) \rb{$\hookrightarrow_P$?} and
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
      \hookrightarrow \machOT'\) are observationally prefixed,
      \(\pi_o(\MPT') \lesssim \pi_o(\machOT')\).

      We write this \(\text{\sc ObsConf}_\contour ~ \MPT\).
      \rb{What about $\machOT$? Ditto for use in the definition of observable stack confidentiality below.}

@@

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109, in pink, are
      not restored; every address that was varied and not changed in one of the
      executions is. Since 108 contains a value that originates in the
      variation, values that depend on it might be anything; after returning
      from {\tt g}, {\tt main} prints 95 in the trace from \(\mach_r\) and
      some undetermined value in the trace from \(\mach_r''\). So observable
      trace confidentiality does not hold.

@@

    \definition{Observable Stack Safety}
      A system enjoys {\em observable stack safety} with regard to a call map
      and a code map if it enjoys both observable \rb{stack?} integrity and observable
      \rb{stack?} confidentiality with regard to them.
